### 2017-09-28 jfm_rapids/jfm-instructions.tex
adding create_flux_pbs_script.py

% This is file JFM2esam.tex
% first release v1.0, 20th October 1996
%       release v1.01, 29th October 1996
%       release v1.1, 25th June 1997
%       release v2.0, 27th July 2004
%       release v3.0, 16th July 2014
%   (based on JFMsampl.tex v1.3 for LaTeX2.09)
% Copyright (C) 1996, 1997, 2014 Cambridge University Press

\documentclass{jfm}
\usepackage{graphicx}
\usepackage{epstopdf, epsfig}

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\shorttitle{Guidelines for authors}
\shortauthor{A. N. Other, H.-C. Smith and J. Q. Public}

\title{Guidelines for authors and submission template}

\author{Alan N. Other\aff{1}
  \corresp{\email{jfm@damtp.cam.ac.uk}},
  H. - C. Smith\aff{1}
 \and J. Q.  Public\aff{2}}

\affiliation{\aff{1}Department of Chemical Engineering, University of America,
Somewhere, IN 12345, USA
\aff{2}Department of Aerospace and Mechanical Engineering, University of
Camford, Academic Street, Camford CF3 5QL, UK}

\begin{document}

\maketitle

\begin{abstract}
An incompressible flow can be an effective mixer of a dye by appropriately advecting the dye to produce small filamentation length scales. It is known that the Batchelor length scale [\cite{Batchelor1959a}] is the theorized lower limit of dye length scales  achievable under turbulent flows. We provide numerical evidence that this limitation may be more general and applies to incompressible flows under mild physical constraints such as fixed enstrophy and energy. We consider local-in-time flow optimization under these constraints with the objective of maximizing mixing rate performance. For enstrophy-bounded optimal flows, we find that the strength of diffusion has no impact on the long-term mixing rate performance. For energy-constrained optimal flows, we find that an increase in the strength of diffusion worsens this performance. We provide analytical lower bounds on mixing rates and length scales achievable under related constraints (pointwise bounded speed and rate-of-strain) by extending the work of \cite{JFM2011} and \cite{Chi-Cheu1996}. 
\end{abstract}

\begin{keywords}
Authors should not enter keywords on the manuscript, as these must be chosen by the author during the online submission process and will then be added during the typesetting process (see http://journals.cambridge.org/data/\linebreak[3]relatedlink/jfm-\linebreak[3]keywords.pdf for the full list)
\end{keywords}

\section{How to submit to the \textbf{\textit{Journal of Fluid Mechanics}}}
Authors must submit using the online submission and peer review system, Scholar One Manuscripts (formerly Manuscript Central) at http://mc.manuscriptcentral.com/jfm. If visiting the site for the first time, users must create

\bibliographystyle{jfm}
% Note the spaces between the initials
\bibliography{manuscript}

\end{document}